---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

*This is a test. let me know what you think.*

# Setting workspace
```{r, echo=FALSE, warning=F}
source("functions.R")
library(tidyverse)

readxl::excel_sheets("REACH_SOM2101_JMCNA_CleanData_v5.xlsx")
cleann <- readxl::read_excel("REACH_SOM2101_JMCNA_CleanData_v5.xlsx", "Clean_Data")
logg <- readxl::read_excel("REACH_SOM2101_JMCNA_CleanData_v5.xlsx", "Cleaning_Log")
raww <- readxl::read_excel("REACH_SOM2101_JMCNA_CleanData_v5.xlsx", "Raw_Data")
dell <- readxl::read_excel("REACH_SOM2101_JMCNA_CleanData_v5.xlsx", "Deletions")
```

# Checking your cleaning log
I am checking your cleaning log against your clean data. 

```{r, echo=FALSE, warning=F}
old_new_values <- mapply(old_new, 
                          cleaning_log = split(logg, row.names(logg)), 
                          variable = "Question", 
                          MoreArgs = list(
                            data_raw = raww,
                            data_clean = cleann,
                            uuid_raw = "uuid",
                            uuid_clean = "uuid",
                            uuid_cleaning_log = "uuid"
                          ),
                          SIMPLIFY = F) %>% do.call(rbind, .)
dd <- logg %>% 
  mutate(binding = paste0(uuid, Question)) %>%
  left_join(old_new_values) %>% 
  mutate(check_raw = `Old Value` == value_raw,
         check_clean = `New Value` == value_clean) %>%
  select(-c(binding, ID, `Follow-up`, Enumerator, Community, `Modified by?`, Notes, check_raw), 
         uuid, Question,`Old Value`, `New Value`, value_raw, value_clean, check_clean, Reason)
```

```{r}
dd$check_clean %>% table(useNA = "ifany")
```

Out of your 25k entries, you have 12k correct, 4k not correct and 10k as NA. 

## NA's
Let's now just focus on the NA
```{r}
dd %>% 
  filter(is.na(check_clean)) %>% 
  group_by(Question) %>%
  tally() %>%
  arrange(desc(n))
```

### How-to read
```{r}
dd %>% head()
```
Question: from your cleaning log   
uuid: uuid   
Old Value: from your cleaning log  
New value: from your cleaning log  
Reason: from your cleaning log  
value_raw : value from raw dataset  
value_clean: value from clean dataset  
check_clean: New Value (from cleaning log) is the same as value_clean (from clean data)  


### ALL 
```{r}
logg %>% filter(Question == "All") %>% mutate(uuid_dup = duplicated(uuid)) 
```
there are 1681 survey deleted, there are 1676 in the deletion log.   
there are 9 uuid duplicated in the cleaning log for deletion  

action: none needed.

### sanitation_features_other: 
```{r}
dd %>% filter(Question == "sanitation_features_other", is.na(check_clean)) 
```
it seems there are 10% of the household who does not have toilets.  

action: to be check with the skip logic 

### main_source_water_other
```{r}
dd %>% filter(Question == "main_source_water_other", is.na(check_clean)) 
dd %>% 
  filter(Question == "main_source_water_other", is.na(check_clean)) %>% 
  select(uuid, `New Value`) %>% 
  left_join(select(cleann, main_source_water, uuid)) 
```
action: Explore what happened, was the "other" recoded?  

### when_arrived_current_location
```{r}
dd %>% filter(Question == "when_arrived_current_location", is.na(check_clean)) %>% 
  View()
```
In your cleaning log, the "New Value" is empty, while there is a value in the clean dataset. No reason written  
action: Explore what happened

### when_leave_place_origin
```{r}
dd %>% filter(Question == "when_leave_place_origin", is.na(check_clean)) %>% 
  View()
```
Same as above.  

### 67"
```{r}
dd %>% filter(Question == "67", is.na(check_clean))   
```
Not sure what 67 variable is for, but same as above. In addition, all of them were turn to 67 while the value were different  
Action: investigate on your side.

### sanitation_facilities_problems_other
```{r}
dd %>% filter(Question == "sanitation_facilities_problems_other", is.na(check_clean)) %>%
  pull(Reason) %>% table()

dd %>% filter(Question == "sanitation_facilities_problems_other", is.na(check_clean), 
              Reason != "other options recoded", 
              Reason != "other options recoded in the choice list") 

dd %>% 
  filter(Question == "sanitation_facilities_problems_other", is.na(check_clean), 
         Reason != "other options recoded", 
         Reason != "other options recoded in the choice list") %>% 
  select(uuid, `New Value`) %>% 
  left_join(select(cleann, sanitation_facilities_problems, uuid)) %>% 
  select(`New Value`, sanitation_facilities_problems) %>% table()
```
Same as for main source of water. Check if recoding happened or should happen. 
Action: explore what happened when recoding other.  


### shelter_issues_other
```{r}
dd %>% 
  filter(Question == "shelter_issues_other", is.na(check_clean)) 

dd %>% 
  filter(Question == "shelter_issues_other", is.na(check_clean))%>% 
  select(uuid, `New Value`) %>% 
  left_join(select(cleann, shelter_issues, uuid)) %>% 
  select(`New Value`, shelter_issues) %>% table()
```
It seems there is "no issues". Should it be removed as you did then? Or removed?
Action: investigate if re-coding correct.

### hh_main_source_income_other
```{r}
dd %>% 
  filter(Question == "hh_main_source_income_other", is.na(check_clean)) %>%
  select(uuid, `Old Value`) %>% 
  left_join(select(cleann, shelter_issues, uuid))

dd %>% 
  filter(Question == "hh_main_source_income_other", is.na(check_clean)) %>% 
  pull(Reason) %>%
  table() 
dd %>% 
  filter(Question == "hh_main_source_income_other", is.na(check_clean)) %>% 
  select(uuid, `Old Value`) %>% 
  left_join(select(cleann, starts_with("hh_main_source_income"), uuid)) 
```
does Clarification from choices and Clarification from translation means re-classfied?  
Action: Check if clarification were correclty re-coded/classified.

### common_type_ids_other
```{r}
dd %>% 
  filter(Question == "common_type_ids_other", is.na(check_clean)) %>%
  select(uuid, `Old Value`, `New Value`) %>% 
  left_join(select(cleann, starts_with("common_type_ids"), uuid)) 
```
Same as the the others "other"  
Action: check recoding and action. 

### General comment
For all others that are removed because of "none", maybe because of missing skip logic during the coding,
do you need to also remove all the questions? or add a "none" in your select mutiple? Or do you want just to report as
that and include the "no-issues" in the denominator?
e.g. do you want to report :
(1) XX % of households reported to have an issue. From the ones reporting an issue, YY% reported as OPTION1 as an issue.
(2) Amongst all the households, XX% reported no issues, YY% reported OPTION1 as an issue.
for (1) you would want to recode the none as missing for all options so that you can use the skiplogic.

## wrong cleaning value. 

```{r}
dd %>% filter(check_clean == F)
dd %>% filter(check_clean == F) %>% nrow()
```

### re-coding FALSE and TRUE to 0/1
```{r}
logg2 <- logg %>% 
  mutate(`New Value` = ifelse(`New Value` == "FALSE", 0, `New Value`), 
         `New Value`= ifelse(`New Value` == "TRUE", 1, `New Value`))

old_new_values2 <- mapply(old_new, 
                          cleaning_log = split(logg2, row.names(logg2)), 
                          variable = "Question", 
                          MoreArgs = list(
                            data_raw = raww,
                            data_clean = cleann,
                            uuid_raw = "uuid",
                            uuid_clean = "uuid",
                            uuid_cleaning_log = "uuid"
                          ),
                          SIMPLIFY = F) %>% do.call(rbind, .)
dd2 <- logg2 %>% 
  mutate(binding = paste0(uuid, Question)) %>%
  left_join(old_new_values2) %>% 
  mutate(check_raw = `Old Value` == value_raw,
         check_clean = `New Value` == value_clean) %>%
  select(-c(binding, ID, `Follow-up`, Enumerator, Community, `Modified by?`, Notes, check_raw), 
         uuid, Question,`Old Value`, `New Value`, value_raw, value_clean, check_clean, Reason)

dd2 %>% filter(check_clean == F) %>% nrow()
```
With changing the T/F we are reducing to 705 wrong values!


```{r}
dd2 %>% filter(check_clean == F)
```

It seems "casuel _labour _Wages _construction _etc" has a typo. 

```{r}
dd2 %>% filter(check_clean == F, 
               `New Value` != "casuel _labour _Wages _construction _etc") %>% nrow()
```
Down to 535. 

```{r}
dd2 %>% filter(check_clean == F, 
               `New Value` != "casuel _labour _Wages _construction _etc")
```

It seems some values are off.   
Action: please check those.

# cleaninginspectoR

```{r}
inspec <- cleaninginspectoR::inspect_all(cleann, "uuid")
```
It seems you have lots of outliers.  

```{r}
inspec %>% 
  filter(issue_type %in% c("log normal distribution outlier", "normal distribution outlier")) %>%
  group_by(variable) %>%
  tally()
```
Action: Please check all those variables for LOW and HIGH values.

# Extra check:

## calculations
I don't have the KOBO so I am just checking the obvious one

### calculation total hh 
```{r}
cleann %>% 
  select(males_0_2y:total_hh) %>% 
  mutate(across(.fns = as.numeric)) %>% 
  mutate(total_cal = rowSums(across(males_0_2y:females_60_older)), 
         check = total_cal == total_hh) %>% 
  pull(check) %>% 
  table(useNA = "ifany")
```
All good   
Action: none needed  


### age_pregnent_female_give_birth
```{r}

cleann$age_pregnent_female_give_birth %>% table(useNA = "ifany")
```

Action: change the 1 year old pregnant woman to NA?   

## checks outliers LOW
I am not sure what does estimate_hh_income refers to but you have very low values all the estimate_hh_income.XXXX have very low values. People earning .20 dollars. See below for anything less than 5(USD?) (0's excluded)
```{r}
cleann %>% 
  select(uuid, starts_with("estimate_hh_income.")) %>% 
  mutate(across(starts_with("estimate_hh_income."), ~(. < 5 & . > 0), .names = "{.col}_less_5"))
```

Action: needed please checked those values. 

## calculation part 2

### hh_children

```{r}
cleann %>% 
  mutate(across(.cols = males_0_2y:females_12_17y, .fns = as.numeric)) %>% 
  mutate(total_cal = rowSums(across(males_0_2y:females_12_17y)), 
         check = total_cal == hh_children) %>% 
  pull(check) %>% 
  table(useNA = "ifany")

cleann %>% 
  mutate(across(.cols = males_0_2y:females_12_17y, .fns = as.numeric)) %>% 
  mutate(total_cal = rowSums(across(males_0_2y:females_12_17y)), 
         check = total_cal == hh_children) %>% 
  filter(check == F) %>%
  select(uuid, males_0_2y:females_12_17y,hh_children, total_cal)
  
```
There are calculations that were not corrected. 
Action: Please ammend

### school_age_children
```{r}
cleann %>% 
  mutate(across(.cols = males_3_5y:females_12_17y, .fns = as.numeric)) %>% 
  mutate(total_cal = rowSums(across(males_3_5y:females_12_17y)), 
         check = total_cal == school_age_children) %>% 
  filter(check == F) %>% 
  select(uuid, males_3_5y:females_12_17y,school_age_children, total_cal)
```
As above. 

### hh_female
The KOBO calculation includes children males and females from 6 to 11. That variable is only used to 
filter on the *pregnant_female_give_birth*. If you are not going to analyse it, just delete it

After checking for only females from 12 to 59, you should turn about 380 *pregnant_female_give_birth* and the 
2 follow-up questions *pregnant_female_give_birth_yes*, *age_pregnent_female_give_birth* to NA. 
```{r}
cleann %>% 
  mutate(across(.cols = c(females_12_17y, females_18_40, females_41_59), .fns = as.numeric)) %>% 
  mutate(check_female = rowSums(across(c(females_12_17y, females_18_40, females_41_59))) > 0) %>% 
  select(check_female, pregnant_female_give_birth) %>% 
  table(useNA = "ifany")
cleann %>% 
  mutate(across(.cols = c(females_12_17y, females_18_40, females_41_59), .fns = as.numeric)) %>% 
  mutate(check_female = rowSums(across(c(females_12_17y, females_18_40, females_41_59))) > 0, 
         to_be_fixed = check_female == F & !is.na(pregnant_female_give_birth) ) %>% 
  filter(to_be_fixed == T) %>%
  select(uuid, check_female, pregnant_female_give_birth, pregnant_female_give_birth_yes, age_pregnent_female_give_birth) 
```

### sum_persons_having_difficulty
You are suming ${seeing_memb}+${hearing_memb}+${walking_climbing_memb}+${remembering_memb}+${selft_care_memb}+${communicating_memb}
and then create a loop from it. If you a person had 2 difficulites, you would have a sum of 2, hence creating a loop for 2 persons and not one. 
Action: how are you going to deal with that? 

### sum_rooms
```{r}
cleann %>% 
  mutate(across(.cols = shelt_count.bedrooms:shelt_count.other_sh, .fns = as.numeric)) %>% 
  mutate(total_cal = rowSums(across(shelt_count.bedrooms:shelt_count.other_sh)), 
         check = total_cal == shelt_count.sum_rooms) %>% 
  filter(check == F) %>% 
  select(uuid, males_3_5y:females_12_17y,school_age_children, total_cal)

```

There 1849 calculation to be corrected.  
Action: please ammend or remove the column if not needed

# list of checks not implemented? 
From your script you shared, you have a list of checks that were not implemented:  
- pregnant_female_give_birth_yes > 2  
- age_pregnent_female_give_birth < 15  
- children_working.permanent_job_child_male > 3  
- children_working.permanent_job_child_female > 3  
- sharing_sanitation_facilities_yes > 14 "how more then 14 hh share a latrine?"  
- how_many_shelters > 8 "how one hh can own 8 or more shelters?"  
- shelter_rooms_note > 3  
- shelt_count.bedrooms > 3  
- shelt_count.living_rooms > 3  
- shelt_count.kitchens > 3  
- shelt_count.toilets	shelt_count.other_sh > 3  
- shelt_count.sum_rooms > 3  

Why were those not checked? 

## Check for shelter rooms above 3
```{r}

cleann %>% 
select(uuid, shelt_count.bedrooms:shelt_count.other_sh) %>% 
mutate(across(shelt_count.bedrooms:shelt_count.other_sh, ~ (.> 3), .names = "{.col}_above_3")) %>% 
summarise(across(ends_with("above_3"), sum))

cleann %>% 
select(uuid, shelt_count.bedrooms:shelt_count.other_sh) %>% 
mutate(across(shelt_count.bedrooms:shelt_count.other_sh, ~ (.> 3), .names = "{.col}_above_3"))

```
## time residency 
I am not sure what this refers to but the number of months is not the conversion of years into months. Please check again those 
variables as they currently do not make sense.  

# HHS 
```{r}
cleann %>% 
  mutate(hhs1 = case_when(how_often_fs_happen == "often_10_plus" ~ 2, 
                         how_often_fs_happen %in% c("rately_1_2", "sometimes_3_10") ~ 1, 
                         is.na(how_often_fs_happen) ~ 0), 
         hhs2 = case_when(often_go_sleep_night_hungry == "often_10_plus" ~ 2, 
                         often_go_sleep_night_hungry %in% c("rately_1_2", "sometimes_3_10") ~ 1, 
                         is.na(often_go_sleep_night_hungry) ~ 0), 
         hhs3 = case_when(often_not_eating_whole_day_night == "often_10_plus" ~ 2, 
                         often_not_eating_whole_day_night %in% c("rately_1_2", "sometimes_3_10") ~ 1, 
                         is.na(often_not_eating_whole_day_night) ~ 0), 
         hhs_score = hhs1 + hhs2 + hhs3, 
         hhs_cat = cut(hhs_score, breaks = c(0,5,7), labels = c("low-none-severe", "extreme"), right = F)) %>%
  select(hhs_cat, district) %>% table(useNA = "ifany") %>% prop.table(2)
  

```
bonndheere, wadajir and yaarqshiid are at the 5% threshold. If any HHS score for a strata is above 5 % we should check
with Matt and Katie.  
Action: Are you going to report as Banadir as a whole or for each city boundary ? (maybe Khada and Dharkenley)

